### 2023-01-24 paper/localpsf.tex
moved text up in the numerics

\subsection{Example 1: Inversion for the basal sliding coefficient in an ice sheet flow problem}

@@

We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady state ice sheet flow \cite{PetraEtAl12}. Briefly, a sheet of ice is flowing down a mountain (see Figure \ref{fig:ice_mountain_mesh}). Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding parameter field, which governs the resistance to sliding along the base of the ice sheet. The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.

\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
		\caption{Bird's eye view}
	\end{subfigure}
	\begin{subfigure}{0.48\textwidth}
		\includegraphics[scale=0.22]{meshHeight_view0_edges_withAxes.png}
		\caption{Top down view}
	\end{subfigure}
	\caption{The ice sheet, discretized by a mesh of tetrahedra. Color indicates the height of the base of the ice sheet (i.e., the mountain topography). The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.}
	\label{fig:ice_mountain_mesh}
\end{figure}

The domain that is filled with ice is denoted by
$\icedomain \subset\mathbb{R}^{3}$. The basal, lateral and top parts of $\partial \icedomain$ are denoted by $\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$, respectively.
The governing equations are the linear incompressible Stokes
equations,

\begin{subequations}\label{Stokeseqn}
	\begin{align}
	-\nabla \cdot \stress
	&=\bodyforce\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
	\nabla\cdot \velocity
	&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
	\stress\normalvec
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
	\velocity \cdot\normalvec =0 \quad\text{and}\quad \tangentop\left(\stress\normalvec+\exp\left(\basalfriction\right)\velocity\right)
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
	\stress\normalvec+\stokesrobincoeff \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
	\end{align}
\end{subequations}
The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field. Here, $\basalfriction$ is the unknown logarithmic
basal sliding field (note that large $\basalfriction$ corresponds to high resistance to sliding) which is defined on the $2$D surface $\Gamma_b$. The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary. 
We employ Glen's flow law \cite{Glen1955}, $\stress= 2\eta
\dot{\strain} -I\pressure$,
which is a constitutive law for ice that relates the stress
tensor, $\stress$, to the strain rate tensor,
$\dot{\strain}= \frac{1}{2}\left(
\nabla\velocity+\nabla\velocity^{\top} \right)$. Here $\eta$ is the effective viscosity and $I$ is the unit matrix. Glen's exponent has been chosen to be one, which provides for a linear Stokes model. Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.

@@

To reconstruct the basal sliding parameter field $\basalfriction$, we
formulate an inverse problem governed by the Stokes
problem~\eqref{Stokeseqn}. In particular this problem is formulated as
a nonlinear least squares optimization problem, whose cost function
consists of a misfit term (between the observations and model output)
and a bi-Laplacian regularization term
following~\cite{IsaacPetraStadlerEtAl15}. The regularization is
centered at the constant function $q_0(x) = 10.5$. To mitigate
boundary effects we use a constant coefficient Robin boundary
condition as in~\cite{Roininen14}. The parameters are chosen so that
the Green's function of the Hessian of the regularization has a
characteristic length of $0.25$ of the domain radius. For the
specific setup, we refer the reader
to~\cite[Section 2.2]{VillaPetraGhattas21}. We choose the regularization
parameter, $\gamma$, that controls the overall strength of the
regularization using Morozov's discrepancy principle. The optimal
value of $\gamma$ is $7.3 \times 10^3$, which is used for all the
numerical results except for Figure~\ref{fig:gamma_sweep}, where we
vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness
of the proposed preconditioner.

%% from here we summarize --------
%The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure \ref{fig:true_beta_u}. We generate synthetic observations by restricting the tangential component of $\velocity_\text{true}$ to the top surface to get
%\begin{equation*}
%\observations_\text{true} = \tangentop \velocity_\text{true} |_{\Gamma_t}.
%\end{equation*}
%We add $1\%$ multiplicative Gaussian noise to all components of the finite element vector, $\bm{\observations}_\text{true}$, corresponding to the function $\observations_\text{true}$, which yields noisy synthetic observations, $\observations_\text{obs}$ which are used for the inversion. That is, 
%\begin{equation*}
%\left(\bm{\observations}_\text{obs}\right)_j := \left(\bm{\observations}_\text{true}\right)_j + 0.01 \cdot \boldsymbol{\xi}_j \cdot \left|\left(\bm{\observations}_\text{true}\right)_j\right|,
%\end{equation*}
%where $j$ ranges over all entries of the vector of coefficients representing the finite element functions $\observations_\text{true}$ and $\observations_\text{obs}$, and $\boldsymbol{\xi}_j$ are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution. We also define the noise function, $\xi$, to be the finite element function with $j$th vector entry $\boldsymbol{\xi}_j$.
%
%The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
%\begin{equation}
%	\label{eq:deterministic_optimization_problem}
%	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
%\end{equation}
%The first term in the objective function is
%\begin{equation*}
%	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
%\end{equation*}
%We call this term the data misfit, because it measures the difference between the observed data and the predicted data based on a candidate parameter $\basalfriction$. Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$. The second term,
%\begin{equation*}
%	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{K}(\basalfriction - \basalfriction_0)\right|^2\,dx,
%\end{equation*}
%is a Bilaplacian regularization term. Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
%\begin{align*}
%-\gamma\Delta u +\delta\,u&=\regrhs\,\,\,\,\text{ in }\Omega,\\
%\gamma \nabla u\cdot\normalvec+\regrobincoeff\,u&=0\,\,\,\text{ on }\partial\Omega.
%\end{align*}
%Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$. Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient \cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation, \eqref{Stokeseqn}.
%In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
%\begin{equation*}
%	\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
%\end{equation*}
%The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$. The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner. The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{K}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain. 
%%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
%

We solve the inverse problem with an inexact Newton preconditioned
conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The (discrete) Newton updates,
$\bm{\searchdir}_k$, are obtained by solving 
%
\begin{equation}
	\label{eq:newton_system}
	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
\end{equation}
%
wherein we choose the initial guess as the discretization of the constant
function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
the discretized gradient, Hessian, and Gauss-Newton Hessian of the
inverse problem cost function, respectively.
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
To ensure positive definiteness of the Hessian we use
$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
subsequent iterations. The Newton iterations are terminated when
$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$. %PCG only requires applying
%$\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors. Each application of
%$\bm{H}$ (or $\bm{H}_\text{gn}$) to a vector requires solving one
%incremental forward and one incremental adjoint Stokes PDE of the form
%\eqref{Stokeseqn}, but with different right hand sides.
\begin{figure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m1_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter1}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m3_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter2}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m4_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter3}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m6_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter4}
	\end{subfigure}
	%\begin{subfigure}{0.05\textwidth}
	%	\includegraphics[scale=0.1]{mColorBar.png}
	%\end{subfigure}
	\caption{The log basal sliding parameter after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right). The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.}
	\label{fig:newton_iterates} 
\end{figure}

\begin{table}
	{\small
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{3pt}
			\renewcommand{\arraystretch}{1.1}
			\begin{tabular}{c| c c c | c c c | c c c}
				&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
				\hline
				Iter & 
				\#CG & \#Stokes & $||\bm{g}||$ & 
				\#CG & \#Stokes & $||\bm{g}||$ & 
				\#CG & \#Stokes & $||\bm{g}||$ \\
				0 &
				1 & 4 & 2.0e+7 &
				1 & 4 & 2.0e+7 &
				1 & 4 & 2.0e+7 \\
				1 &
				2 & 6 & 5.8e+6 &
				2 & 6 & 7.9e+6 &
				2 & 6 & 5.8e+6 \\
				2 &
				4 & 10 & 2.4e+6 &
				5 & 12 & 3.9e+6 &
				4 & 10 & 2.4e+6 \\
				3 &
				2 & 6+22 & 5.8e+5 &
				13 & 28 & 1.6e+6 &
				13 & 28 & 5.8e+5 \\
				4 &
				5 & 12 & 5.6e+4 &
				42 & 86 & 4.8e+5 &
				33 & 68 & 9.1e+4 \\
				5 &
				10 & 22 & 3.5e+3 &
				84 & 170 & 7.7e+4 &
				57 & 116 & 6.2e+3 \\
				6 &
				14 & 30 & 2.7e+1 &
				125 & 252 & 5.0e+3 &
				76 & 154 & 1.1e+2 \\
				7 &
				0 & 2 & 4.1e-2 &
				194 & 390 & 7.9e+1 &
				117 & 236 & 2.3e-1 \\
				8 & 
				--- & --- & --- &
				0 & 2 & 1.6e-1 &
				0 & 2 & 4.1e-2 \\
				\hline
				Total & 
				38 & 114 & --- &
				466 & 950 & --- &
				303 & 624 & --- \\
			\end{tabular}
			\endgroup
		\end{center}
	}
	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG. Preconditioners shown are our method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE). Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$. Columns titled $||\bm{g}||$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}_k$. Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration. This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system. In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner. }
	\label{tab:newton_convergence_table}
\end{table}

@@

\subsection{Example 2: Inversion for the initial condition in an advection-diffusion problem}

>>>>

We use our method to approximate the Newton (or Gauss-Newton) Hessian
in an inverse problem governed by a PDE which models steady state ice
sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and transport
and flow of a contaminant~\cite{PetraStadler11}
(Section~\ref{sec:adv}). To reconstruct the unknown parameters, we
formulate an inverse problem governed by the corresponding PDEs. In
particular this problem is formulated as a nonlinear least squares
optimization problem, whose cost function consists of a misfit term
(between the observations and model output) and a bi-Laplacian
regularization term following~\cite{VillaPetraGhattas21}. The
regularization is centered at a constant function $q_0(x)$. To
mitigate boundary effects
%For stokes q_0(x) = 10.5$%
we use a constant coefficient Robin boundary condition as
in~\cite{Roininen14}. The parameters for the regularization are chosen
so that the Green's function of the Hessian of the regularization has
a characteristic length of $0.25$ of the domain radius. For the
specific setup, we refer the reader to~\cite[Section
  2.2]{VillaPetraGhattas21}. For both examples, we choose the
regularization parameter, $\gamma$, that controls the overall strength
of the regularization using Morozov's discrepancy
principle~\cite{Vogel02}. The optimal value of $\gamma$ is used for
%  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
all the numerical results except for the studies where study how
changing $\gamma$ impacts the effectiveness of the proposed
preconditioner.

We solve the inverse problem with an inexact Newton preconditioned
conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The (discrete) Newton updates,
$\bm{\searchdir}_k$, are obtained by solving 
%
\begin{equation}
	\label{eq:newton_system}
	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
\end{equation}
%
wherein we choose the initial guess as the discretization of the constant
function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
the discretized gradient, Hessian, and Gauss-Newton Hessian of the
inverse problem cost function, respectively.
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
To ensure positive definiteness of the Hessian we use
$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
subsequent iterations. The Newton iterations are terminated when
$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$.

\subsection{Example 1: Inversion for the basal sliding coefficient in an ice sheet flow problem}
\label{sec:ice}

@@

For this example, we consider a sheet of ice flowing down a mountain (see Figure \ref{fig:ice_mountain_mesh}). Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding Robin coefficient field, which governs the resistance to sliding along the base of the ice sheet. %The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
		\caption{Bird's eye view}
	\end{subfigure}
	\begin{subfigure}{0.48\textwidth}
		\includegraphics[scale=0.22]{meshHeight_view0_edges_withAxes.png}
		\caption{Top down view}
	\end{subfigure}
	\caption{The ice sheet, discretized by a mesh of tetrahedra. Color indicates the height of the base of the ice sheet (i.e., the mountain topography). The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.}
	\label{fig:ice_mountain_mesh}
\end{figure} 
The domain that is filled with ice is denoted by
$\icedomain \subset\mathbb{R}^{3}$. The basal, lateral and top parts of $\partial \icedomain$ are denoted by $\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$, respectively.
The governing equations are the linear incompressible Stokes
equations,

\begin{subequations}\label{Stokeseqn}
	\begin{align}
	%-\nabla \cdot \stress(v, p)
	%&=\bodyforce \,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
	%\nabla\cdot \velocity
	  %&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
          -\nabla \cdot \stress(v, p)=\bodyforce \text{ and }
          \nabla\cdot \velocity &=0 \quad \text{ in }\icedomain, \label{Stokeseqn:1}\\
	\stress(v, p)\normalvec
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
	\velocity \cdot\normalvec =0 \text{ and } \tangentop\left(\stress(v, p)\normalvec+\exp\left(\basalfriction\right)\velocity\right)
	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
	\stress(v, p)\normalvec+\stokesrobincoeff \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
	\end{align}
\end{subequations}
The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field. Here, $\basalfriction$ is the unknown logarithmic
basal sliding field (note that large $\basalfriction$ corresponds to high resistance to sliding) which is defined on the $2$D surface $\Gamma_b$. The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary. 
We employ Glen's flow law \cite{Glen1955}, $\stress(v, p)= 2\eta
\dot{\strain}(v) -I\pressure$,
which is a constitutive law for ice that relates the stress
tensor, $\stress$, to the strain rate tensor,
$\dot{\strain}(v)= \frac{1}{2}\left(
\nabla\velocity+\nabla\velocity^{\top} \right)$. Here $\eta$ is the effective viscosity and $I$ is the unit matrix. Glen's exponent has been chosen to be one, which provides for a linear Stokes model. Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.

@@

%aaa

%% from here we summarize --------
%The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure \ref{fig:true_beta_u}. We generate synthetic observations by restricting the tangential component of $\velocity_\text{true}$ to the top surface to get
%\observations_\text{true} = \tangentop \velocity_\text{true} |_{\Gamma_t}.
%We add $1\%$ multiplicative Gaussian noise to all components of the finite element vector, $\bm{\observations}_\text{true}$, corresponding to the function $\observations_\text{true}$, which yields noisy synthetic observations, $\observations_\text{obs}$ which are used for the inversion. That is, 
%\left(\bm{\observations}_\text{obs}\right)_j := \left(\bm{\observations}_\text{true}\right)_j + 0.01 \cdot \boldsymbol{\xi}_j \cdot \left|\left(\bm{\observations}_\text{true}\right)_j\right|,
%where $j$ ranges over all entries of the vector of coefficients representing the finite element functions $\observations_\text{true}$ and $\observations_\text{obs}$, and $\boldsymbol{\xi}_j$ are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution. We also define the noise function, $\xi$, to be the finite element function with $j$th vector entry $\boldsymbol{\xi}_j$.
%The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
%	\label{eq:deterministic_optimization_problem}
%	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
%The first term in the objective function is
%	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
%We call this term the data misfit, because it measures the difference between the observed data and the predicted data based on a candidate parameter $\basalfriction$. Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$. The second term,
%	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{K}(\basalfriction - \basalfriction_0)\right|^2\,dx,
%is a Bilaplacian regularization term. Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
%\begin{align*}
%-\gamma\Delta u +\delta\,u&=\regrhs\,\,\,\,\text{ in }\Omega,\\
%\gamma \nabla u\cdot\normalvec+\regrobincoeff\,u&=0\,\,\,\text{ on }\partial\Omega.
%\end{align*}
%Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$. Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient \cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation, \eqref{Stokeseqn}.
%In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
%	\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
%The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$. The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner. The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{K}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain. 
%%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
%PCG only requires applying
%$\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors. Each application of
%$\bm{H}$ (or $\bm{H}_\text{gn}$) to a vector requires solving one
%incremental forward and one incremental adjoint Stokes PDE of the form
%\eqref{Stokeseqn}, but with different right hand sides.
\begin{figure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m1_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter1}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m3_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter2}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m4_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter3}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m6_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter4}
	\end{subfigure}
	%\begin{subfigure}{0.05\textwidth}
	%	\includegraphics[scale=0.1]{mColorBar.png}
	%\end{subfigure}
	\caption{The log basal sliding parameter after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right). The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.}
	\label{fig:newton_iterates} 
\end{figure}

\begin{table}
	{\small
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{3pt}
			\renewcommand{\arraystretch}{1.1}
			\begin{tabular}{c| c c c | c c c | c c c}
				&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
				\hline
				Iter & 
				\#CG & \#Stokes & $||\bm{g}||$ & 
				\#CG & \#Stokes & $||\bm{g}||$ & 
				\#CG & \#Stokes & $||\bm{g}||$ \\
				0 &
				1 & 4 & 2.0e+7 &
				1 & 4 & 2.0e+7 &
				1 & 4 & 2.0e+7 \\
				1 &
				2 & 6 & 5.8e+6 &
				2 & 6 & 7.9e+6 &
				2 & 6 & 5.8e+6 \\
				2 &
				4 & 10 & 2.4e+6 &
				5 & 12 & 3.9e+6 &
				4 & 10 & 2.4e+6 \\
				3 &
				2 & 6+22 & 5.8e+5 &
				13 & 28 & 1.6e+6 &
				13 & 28 & 5.8e+5 \\
				4 &
				5 & 12 & 5.6e+4 &
				42 & 86 & 4.8e+5 &
				33 & 68 & 9.1e+4 \\
				5 &
				10 & 22 & 3.5e+3 &
				84 & 170 & 7.7e+4 &
				57 & 116 & 6.2e+3 \\
				6 &
				14 & 30 & 2.7e+1 &
				125 & 252 & 5.0e+3 &
				76 & 154 & 1.1e+2 \\
				7 &
				0 & 2 & 4.1e-2 &
				194 & 390 & 7.9e+1 &
				117 & 236 & 2.3e-1 \\
				8 & 
				--- & --- & --- &
				0 & 2 & 1.6e-1 &
				0 & 2 & 4.1e-2 \\
				\hline
				Total & 
				38 & 114 & --- &
				466 & 950 & --- &
				303 & 624 & --- \\
			\end{tabular}
			\endgroup
		\end{center}
	}
	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG. Preconditioners shown are our method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE). Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$. Columns titled $||\bm{g}||$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}_k$. Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration. This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system. In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner.}
	\label{tab:newton_convergence_table}
\end{table}

@@

\subsection{Example 2: Inversion for the initial condition in an advection-diffusion problem}
\label{sec:adv}
